---
title: "R Day 2"
author: "Ziying Wang"
date: "9/18/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
here::here("data", "gapminder5.csv")
gapminder <- read.csv(here::here("data", "gapminder5.csv"))

```

##Converting factor to characters

```{r}
gapminder$country <- as.character(gapminder$country)
gapminder$continent <- as.character(gapminder$continent)
str(gapminder)

```

```{r}
obs <- 1:nrow(gapminder)
for (i in obs){
  gapminder[i, "gdp"] <- gapminder[i,"pop"]*gapminder[i, "gdpPercap"]
  
}
```

```{r}
for(i in obs){
  gapminder[i, "log_gdpPercap"] <- log(gapminder[i, "gdpPercap"]) 
  gapminder[i, "log_pop"] <- log(gapminder[i, "pop"])
}
```

```{r}
head(gapminder)
```


```{r}
years <- unique(gapminder$year)
for (i in years){
  mean_le <- mean(gapminder$lifeExp[gapminder$year == i], na.rm = T)
  print(paste0(i, ": ", mean_le))
}
```
Which continent has the highest mean life expectancy
```{r}
continents <- unique(gapminder$continent)

for (i in continents){
  mean_le <- mean(gapminder$lifeExp[gapminder$continent == i], na.rm = T)
  print(paste0(i, ": ", mean_le))
}

```

What is the mean life expectancy for each continent for each year?

```{r}
for (i in continents){
  print(paste0("continent: ", i))
  for (j in years){
    mean_le <- mean(gapminder$lifeExp[gapminder$continent == i & gapminder$year == j], na.rm = T)
    print(paste0(j, ": ", mean_le))
    
  }
  
}
```
Has the gap in life expectancy between countries on different continents narrowed over time?
```{r}
for (i in years){
  print(paste0("year: ",i))
  for(j in continents){
    sd_le <- sd(gapminder$lifeExp[gapminder$continent == j & gapminder$year ==i ], na.rm = T)
    print(paste0(j, ": ", sd_le))
  }
}
```
```{r}
vars <- gapminder[,c("lifeExp", "pop", "gdpPercap")]
apply(vars, 2, mean)

```
```{r}
lapply(gapminder, mean)
```
```{r}
sapply(years, function(x) mean(gapminder$lifeExp[ gapminder$year == x ]))
```
What is the standard deviation for life expectancy for each year between 1987 and 2002 (inclusive)?
```{r}
i <- 1987
while(i <= 2002){
  sd_lf <- sd(gapminder$lifeExp[gapminder$year == i])
  print(paste0(i , ": ", sd_lf))
  i <- i +5
}
```

```{r}

set.seed(10)
random_year <- sample(years, 1)

if (random_year > 1977) {
  print(random_year)
}
```

Write a for loop that reports the mean population for years greater than or equal to 1987. Make sure the loop prints a message if the condition is not met!

```{r}
threshold<-1987

for (i in unique(gapminder$year)){
  tmp <- mean(gapminder$lifeExp[gapminder$year==i])
  if (i < threshold){
    print(paste("Mean life expectancy in", i, "is less than the threshold."))
  }else{
    print(paste("Mean life expectancy in", i, "is greater than the threshold. The mean pop is",tmp))
  }

}

```

```{r}
get_value <- function(df, variable = "continent"){
  val <- unique(df[variable])
  print(paste0(variable, ": ", val))
}
```

```{r}
get_mean_sd <- function(df, variable, country){
  var <- df[[variable]][df$country == country]
  m_le <- mean(var)
  sd_le <- sd(var)
  cat("Country", country, "\nMean life Expectancy", m_le, "\nSD life Expectancy", sd_le)
}

get_mean_sd(gapminder, "lifeExp", "Bulgaria")
```
Write a function that reports the mean, median, minimum, and maximum for life expectancy for a continent in gapminder
```{r}
get_inform <- function(df, variable, continent){
  var <- df[[variable]][df$continent == continent]
  m_le <- mean(var)
  medium <- median(var)
  minimum <- min(var)
  maximum <- max(var)
  cat("Continent", continent, "\nMean life Expectancy", m_le, "\nMedian life Expectancy", medium, "\nMinimum life Expectancy", minimum, "\nMaximum life Expectancy", maximum)
}

get_inform(gapminder,"lifeExp", "Asia")
```
